Contextual Identity: A Framework for Usable Privacy

### 2012-10-13 paper.tex
audited uses of p as identity

Helen Nissenbaum has long argued that privacy violations come not from the
simple sharing of personal information, but rather from sharing that
information in a way that violates social norms, or in the wrong
context~\cite{nissenbaum}. The desire for spontaneous, positive human
interaction often leads to sharing personal information. For example, take
PatientsLikeMe~\cite{patientslikeme}. This site functions as a support group
and as a place for people to share health information, such as drug regimens,
treatment history, side-effects, and more. Though this information is
traditionally considered extremely sensitive, the users of this site clearly
value the benefits of sharing (longitudinal graphs of their own health history,
building community with people who share their disease). Further, sharing
personal information does not mean that the users don't value privacy.  Selling
user data originally intended for one purpose for another (e.g., to insurance
companies to discover uninsurable patients) would be a gross violation of
users' privacy expectations.

@@

In offline environments, managing contextual identities is more intuitive than
in online environments. There are cues to help you determine where you are,
who your intended audience is, how many people will overhear you, and how
likely your information is to be re-broadcast in a different context. It is
also true that in offline environments, humans don't have perfect memories and
will eventually what forget.  However, with the advent of increasingly vast
searching, indexing, and archiving capabilities, one cannot rely on forgetting
in online enviroments~\cite{delete}. Thus, managing contextual identities 
online has become increasingly difficult and fraught with mistakes.

@@

We provide a simple formalization of the notion of contextual identity, use it as a
framework to understand existing privacy violations, and analyze mitigations of
those violations.

@@

Many authors have noted how re-broadcasting information information out of
context and making information discoverable can lead to user distress (even if
the information was previously public, but not easily
discoverable)~\cite{boyd1,chew,nissenbaum}.

Barth et al.~provide a formal model for contextual integrity of user
data~\cite{barth}. The contextual integrity model focuses on the flow of
personal information in terms of knowledge states held by agents rather than on
user identity, and works well for modeling privacy legislation and policy. In
contrast, this work's model provides a simple framework to understand users' ability
to represent which identity they want to project depending on context, and works
well for modeling how software features affect users' ability to assert choice
and control.

@@

\begin{definition}
  Given:
  \begin{itemize}
    \item $R$, a set of actors and their relationships to the user
    \item $T$, a set of tasks and their side-effects
    \item $p$, an identity projected by a user or perceived by actors in $R$
  \end{itemize}
\medskip

  We define a Contextual Identity $I = \{R,T,p\}$, where a user performs tasks
$T$ in view of or interaction with actors in $R$, while representing identity
$p$.

\end{definition}

A user may have multiple contextual identities $\{I\}$. These identities change
over time, as users form and dissolve relationships, and focus on different
sets of tasks. The user's goal is to accomplish tasks in $T$,
interact with those in $R$ and claim to be $p$  while being able to predict and
assert control over as much of $I$ as possible.

Our definition focuses on relationships because most privacy concerns stem
from causing an unexpected exchange of data with people the users
knows~\cite{wang}. Tasks are key because users are task-oriented, and it is
easy to map user actions (page navigation, sending email, posting, filling
forms) to tasks.

\begin{definition}
  A Contextual Identity Violation occurs when
parameters $R,T,p$ are controlled by others.  When a user's tasks are shared
with the wrong set $R$, the wrong projected identifier $p$ is employed, or a
user is perceived as performing tasks $T$ they do not anticipate, a
violation of the integrity of Contextual Identity occurs.

\end{definition}

@@

\subsection{Expiring contextual identities}
Oftentimes a user may have a long-term task, such as wedding planning or house
buying, that requires constructing a contextual identity. The tasks associated
with this identity have an externally imposed deadline and may have long-lived
side-effects, such as long-term cookies that reveal sensitive information
(e.g., the neighborhood of the house purchase, income information derived from
purchases, other demographic data). For contextual identities that have
outlived their usefulness, users would benefit from destruction of side-effects
in $T$ such as cookie data.

@@

\subsubsection{Remembering to Forget}
Humans don't have perfect memories, and perhaps social
networks should not either~\cite{delete}. Many users already engage in manual auditing and
deletion of old posts~\cite{fbtips2}. A better solution would be to build tools
that let the user manage this more easily through an API: both Twitter and Facebook provide
APIs for deletion\footnote{At the time of this writing, Google Plus offers a
read-only API}. One way to reduce the likelihood of linked contextual
identities or assert better control over $p$ would be to provide applications that expire posts after some time.
Deletion ane expiration also reduce $R$ (in the case of public posts) and $T$.
%This would help users better contain the whole context ($R,T$) by segmenting
%one large audience and one large set of tasks into smaller temporal partitions.

@@

The Google Sharing project allows users of Google services to confound the way
Google identifies them~\cite{googlesharing}.  The effect is that users employ a
contextual identity with a shared $p$.  These users
employ it only for Google sites and can accomplish tasks
while associating all their actions with a random $p$ also associated
with many other users.  This devalues $p$ and makes it difficult to correlate
tasks with a particular user.

>>>>

Helen Nissenbaum has long argued that privacy violations come not from the
simple sharing of personal information, but rather from sharing that
information in a way that violates social norms, or in the wrong
context~\cite{nissenbaum}. The desire for spontaneous, positive human
interaction often leads to sharing personal information. For example, take
PatientsLikeMe~\cite{patientslikeme}. This site functions as a support group
and as a place for people to share health information, such as drug regimens,
treatment history, side-effects, and more. Though this information is
traditionally considered extremely sensitive, the users of this site clearly
value the benefits of sharing: longitudinal graphs of their own health history,
building community with people who share their disease. Further, sharing
personal information does not mean that the users don't value
privacy~\cite{boyd2}. Selling user data originally intended for one purpose for
another (e.g., to insurance companies to discover uninsurable patients) would
be a gross violation of users' privacy expectations.

@@

In offline environments, managing contextual identities is more intuitive than
in online environments. There are cues to help you determine where you are,
who your intended audience is, how many people will overhear you, and how
likely your information is to be re-broadcast in a different context. It is
also true that in offline environments, humans don't have perfect memories and
will eventually forget.  However, with the advent of increasingly vast
searching, indexing, and archiving capabilities, one cannot rely on forgetting
in online enviroments~\cite{delete}. Thus, managing contextual identities
online has become increasingly difficult and fraught with mistakes.

@@

We provide a simple formalization of the notion of contextual identity, use it
as a framework to understand existing privacy violations, and analyze
mitigations for those violations.

@@

Many authors have noted how re-broadcasting information information out of
context and making information discoverable can lead to user distress, even if
the information was previously public, but not easily
discoverable~\cite{boyd1,chew,nissenbaum}.

Barth et al.~provide a formal model for contextual integrity of user
data~\cite{barth}. The contextual integrity model focuses on the flow of
personal information in terms of knowledge states held by agents rather than on
user identity, and works well for modeling privacy legislation and policy. In
contrast, this work's model provides a simple framework to understand users'
ability to represent which contextual identity they want to project, and works
well for modeling how software features affect users' ability to assert choice
and control.

@@

Suppose:
  \begin{itemize}
    \item $R$ is a set of actors and their relationships to the user. $R$ may include people that the user knows, service providers, or network operators.
    \item $T$ is a set of tasks and their side-effects. Examples of tasks
include page navigation, posting to social networks and searching. Examples of
side-effects include cookies, geolocation data and browser history.
    \item $p$ is an impression that the user wishes to project, or that actors
in $R$ perceive. $p$ encompasses such prosaic information as email identifiers
or session identifiers, or such sentiment as ``Alice is awesome!''
  \end{itemize}

\begin{definition} We define a contextual identity $I = \{R,T,p\}$, where a user performs tasks
$T$ in view of or interaction with actors in $R$, while projecting impression
$p$.
\end{definition}

A user may have multiple contextual identities $\{I\}$. These identities change
over time, as users form and dissolve relationships, and focus on different
sets of tasks. The user's goal is to accomplish tasks in $T$,
interact with those in $R$ and project $p$ while being able to predict and
assert control over as much of $I$ as possible.

Our definition focuses on relationships because most privacy concerns stem
from causing an unexpected exchange of data with people the user
knows~\cite{wang}. Tasks are key because users are task-oriented, and it is
easy to map user actions (page navigation, sending email, posting, filling
forms) to tasks.

\begin{definition}
A contextual identity violation occurs when the user cannot predict or control
parameters $R,T,p$.
\end{definition}
When a user's tasks are shared with the wrong set $R$, the wrong projected
identifier $p$ is employed, or a user is perceived as performing tasks $T$ they
do not anticipate, a violation of the integrity of contextual identity occurs.

\begin{definition}
Linkage of contextual identities occurs when a set of $\{I\}$ is associated
with the same user.
\end{definition}

When this linkage happens against the will or without the knowledge of the
user, a violation of the integrity of contextual identity occurs.

@@

\subsection{Expiring contextual identities}
Oftentimes a user may have a long-term task, such as wedding planning or house
buying, that requires constructing a contextual identity. The tasks associated
with this identity have an externally imposed end date and may have long-lived
side-effects, such as long-term cookies that reveal sensitive information
(e.g., the neighborhood of the house purchase, income information derived from
purchases, other demographic data). For contextual identities that have
outlived their usefulness, users would benefit from destruction of side-effects
in $T$ such as cookie data.

@@

\subsubsection{Remembering to Forget}
Humans don't have perfect memories, and perhaps social
networks should not either~\cite{delete}. Many users already engage in manual auditing and
deletion of old posts~\cite{fbtips2}. A better solution would be to build tools
that let the user manage this more easily through an API: both Twitter and Facebook provide
APIs for deletion\footnote{At the time of this writing, Google Plus offers a
read-only API}. One way to reduce the likelihood of linked contextual
identities or assert better control over $p$ would be to provide applications that expire posts after some time.
Deletion and expiration also reduce $R$ especially in the case of public posts
and $T$.
%This would help users better contain the whole context ($R,T$) by segmenting
%one large audience and one large set of tasks into smaller temporal partitions.

@@

The Google Sharing project allows users of Google services to confound the way
Google identifies them~\cite{googlesharing}. In this case $R$ contains Google
and the Google Sharing proxy. The effect is that users employ a contextual
identity with a shared $p$. These users employ it only for Google sites not
requiring login and can accomplish tasks while associating all their actions
with a random $p$ also associated with many other users.  This devalues $p$ and
makes it difficult for Google to correlate tasks with a particular user.
